---
title: 'Motor Trend Analysis: regmods-011'
author: "mjrobichaud"
date: "February 21, 2015"
output: pdf_document
---

## Executive Summary
From the analysis detailed below, it can be determined that transmission has a significant effect on fuel economy, with manual transmissions having a 7.24 mpg improvement on average.

## Loading Data and Exploratory Analysis

First, we need to load the mtcars analysis
```{r}
data(mtcars)
names(mtcars)
```

Using ?mtcars, we can get a description for the dataset. The relevant factor for this analysis is the transmission

[, 9]  am	Transmission (0 = automatic, 1 = manual)

Specifically, we're interested in knowing the transmission type and its effect on fuel economy (MPG)

See appendix for boxplot of mpg range for transmission type.

## Answering The First Question
“Is an automatic or manual transmission better for MPG”

To answer this question, we first look at the mean fuel economy aggregated by transmission type:
```{r}
aggregate(mpg~am, data = mtcars, mean)
```

This shows that manual transmissions, on average, have better fuel economny than automatic transmissions.

A simple t-test will allow us to reject the null hypothesis, that automatic transmissions average better fuel economy than manual transmissions:
```{r}
m_trans_data <- mtcars[mtcars$am == 1,]
a_trans_data <- mtcars[mtcars$am == 0,]

t.test(m_trans_data$mpg, a_trans_data$mpg)
```
A p-value of 0.001374 allows us to reject this null hypothesis.

## Answering The Second Question
"Quantify the MPG difference between automatic and manual transmissions"

First, we can create a regression model:
```{r}
model <- lm(mpg~am, data = mtcars)
summary(model)
```
The summary shows that the adjusted R-squared value is 0.3385, meaning that transmission type only accounts for 33.9% of the variance. The difference is 7.24 mpg.

\pagebreak

## Appendix
```{r}
boxplot(mpg~am,
        data = mtcars,
        main = "Fuel Economy by Transmission Type",
        xlab = "Transmission Type",
        ylab = "Fuel Economy (mpg)",
        names = c("Automatic", "Manual"))
```

